Airbnb Exploration by Fernando Troeman
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(memisc)
library(dplyr)
library(ggplot2)
library(GGally)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load the Data
airbnb_full = read.csv('Airbnb NY Dataset.csv')
airbnb = subset(airbnb_full, select=c('neighbourhood_group_cleansed', 
                                      'room_type', 'accommodates', 'bedrooms', 
                                      'price', 'guests_included', 
                                      'availability_365', 'number_of_reviews', 
                                      'review_scores_rating', 
                                      'cancellation_policy', 
                                      'reviews_per_month'))
colnames(airbnb)[1] <- 'borough'
colnames(airbnb)[9] <- 'review_score'
```

The dataset we are exploring contains Airbnb listings data for New York City, 
updated as of 02 October, 2017. It provides listing details such as the number 
of bed and bathrooms, reviews and ratings, and price per night. It contains a 
total of 44,317 Airbnb listings with up to 96 features for each listing. 
However, we will only be conducting a deep-dive into 11 of these features.

# Univariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
ggplot(aes(x = reorder(borough, borough, length)), data = airbnb) + 
  geom_histogram(stat="count") +
  xlab("Boroughs") +
  ylab("Number of Listings")
```

Most of the Airbnb listings in New York City are concentrated in the boroughs 
of Manhattan and Brooklyn, an expected observation given that Airbnb is mainly 
utilized by travellers and tourists who would generally find these two boroughs 
more attractive. We can also reasonably deduce that the availability of listings
are dictated by demand rather than supply, as the number of listings in each 
borough are incommensurate with its population. 2016 figures show that in terms 
of population, Queens (2.3 million) is second only to Brooklyn (2.6 million). 
Furthermore, the Bronx (1.5 million) does not lag far behind Manhattan 
(1.6 million). 

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=reorder(room_type, room_type, length)), data=airbnb) +
  geom_histogram(stat="count") +
  xlab("Room Type") +
  ylab("Number of Listings")
```

The histogram above tells us that listings offering an entire home/apartment or 
private rooms are by far more popular than listings offering shared rooms. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=cancellation_policy), data=subset(airbnb, cancellation_policy 
                                               %in% c("flexible", "moderate", 
                                                      "strict"))) +
  geom_histogram(stat="count") +
  xlab("Cancellation Policy") +
  ylab("Number of Listings")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(airbnb$cancellation_policy)
```

In the histogram above, we filtered out listings with obscure cancellation 
policies - a total of 17 listings split between 3 categories. For cancellation 
policies classified as Flexible, Moderate or Strict, we see that most listings 
have a Strict cancellation policy.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=price), data=airbnb) +
  geom_histogram(binwidth=100) +
  xlab("Price") +
  ylab("Number of Listings")
```

A plot of price presents us with long-tailed data as there are a few ultra 
high-end rentals causing the histogram to be skewed greatly. A logarithmic 
transformation of the data, as shown below, provides a much clearer picture of 
the price distribution. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=price), data=airbnb) +
  geom_histogram(binwidth=0.05) +
  scale_x_log10() +
  xlab("Price") +
  ylab("Number of Listings")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(airbnb$price)
```

We see that 50% of all prices are between \$70 and \$175. More in-depth 
multivariate analysis on prices will be conducted, as we seek to examine how 
price might vary across a number of factors, such as the number of rooms, 
location, and review scores.


```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=accommodates), data=airbnb) +
  geom_histogram(stat="count") + 
  scale_x_continuous(breaks = seq(1,16,1)) +
  xlab("Number of People Listing Accommodates") +
  ylab("Number of Listings")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=guests_included), data=airbnb) +
  geom_histogram(stat="count") + 
  scale_x_continuous(breaks = seq(1,16,1)) +
  xlab("Number of Guests Included in Listing Reservation") +
  ylab("Number of Listings")
```

From the plots above, we see that most listings accommodate up to 2 people. 
However, we also see that, usually, only 1 guest is included in a reservation 
for any listing. As there are often extra charges for additional guests, this 
is an indication of potential 'hidden costs'.

Moving on, we take a look at the review scores assigned to listings by Airbnb 
users.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=review_score), data=subset(airbnb, 
                                                !is.na(review_score))) +
  geom_histogram(stat="count") +
  xlab("Review Score Ratings") +
  ylab("Number of Listings")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(subset(airbnb, !is.na(review_score))$review_score)
```

An initial overview of review score ratings (for listings with at least 1 
review) tells us that most listings have a score rating of 100. The median 
review score is 96.00 while the mean is 93.49.

However, if we establish a strict criteria and only look at listings with 
more than 10 reviews, a more realistic picture emerges and the proportion of 
listings with a score of 100 dips drastically.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=review_score), data=subset(airbnb, 
                                                !is.na(review_score) & 
                                                  number_of_reviews > 10)) +
  geom_histogram(stat="count") +
  xlab("Review Score Ratings (for Listings with >10 Reviews)") +
  ylab("Number of Listings")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(subset(airbnb, !is.na(review_score) & 
                 number_of_reviews > 10)$review_score)
```

For listings with more than 10 reviews, the median review score drops to 94.00 
but the mean review score remains the same at 93.49. The resultant plot and the 
unchanged mean suggest that the criteria implemented not only removed many 
listings with a perfect review score, but also listings with anomalously low 
review scores. Later on, we will continue to further investigate the 
relationship between price and review scores, as one might expect listing owners
to charge higher prices for well-received listings.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=reviews_per_month), data=subset(airbnb, 
                                             !is.na(reviews_per_month))) +
  geom_histogram(binwidth=0.05) +
  scale_x_log10() +
  xlab('Reviews per Month') +
  ylab('Number of Listings')
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(airbnb$reviews_per_month)
```

On average, each listing receives 1.446 reviews per month. The median number of 
reviews per month is 0.870. The number of reviews each listing received per 
month is useful as it helps us get a sense of the frequency of bookings for any 
particular listing - Airbnb has stated that 70% of all stays end up with a 
review.

Next, we will have a look at the distribution of listings that have been active 
(available to book for at least 1 day) within the past year.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=availability_365), data=subset(airbnb, availability_365 != 0)) +
  geom_histogram(stat = "count") +
  xlab('Number of Days Available') +
  ylab('Number of Listings')
```

49% of Airbnb listings located in New York City have been available for a total 
of 0 days (out of the previous 365 days ending on 02 October, 2017) - taking 
those listings into account while plotting listing availability would present a 
highly skewed figure. As a result, we generated a plot of listing availability 
for listings that have been available for booking for at least one day within 
the past year. We see that the modal availability is 365 days.

# Univariate Analysis

### What is the structure of your dataset?

The dataset consists of 44,317 Airbnb listings in New York City. There are a 
total of 96 features, but we will only be taking a deep dive into 11 of those 
features.

These 11 features are as follows:
  1. Borough
  2. Room Type
  3. Number of People Listing Accommodates
  4. Number of Bedrooms
  5. Price
  6. Number of Guests Included
  7. Number of Days Available (Within The Past Year)
  8. Number of Reviews
  9. Review Score
  10. Cancellation Policy
  11. Number of Reviews per Month

### What is/are the main feature(s) of interest in your dataset?

This exploratory analysis primarily aims to examine the relationship between 
price and the various features of any particular listing. Factors that are 
expected to have a big impact on price include location and the number of 
bedrooms.

In the process of identifying relevant factors contributing to the price of a 
listing, I am hopeful of discovering some peripheral factors that have an 
unexpectedly large impact on the price of a listing.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Some other factors that could potentially have a significant influence on the 
price of a listing include the availability of the listing, review scores, 
cancellation policy and the frequency of bookings at the listing.

### Did you create any new variables from existing variables in the dataset?

No.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

An initial investigation of the data revealed that as much as 49% of the 
listings had not been available for booking in the 365 days ending on 
02 October, 2017. This prompts caution when conducting analysis relevant to 
the availability of Airbnb listings, as some of the data pertaining to inactive
listings might be outdated.

The data was also occasionally adjusted to group the listings by certain 
features and then to extract statistics and insights from these groups.

# Bivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
ggpairs(subset(airbnb, select=-c(room_type, number_of_reviews)))
```

Firstly, we would like to test our hypothesis that location and the number of 
bedrooms are the biggest determinants of the price of an Airbnb listing.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=borough, y=price), data=airbnb) +
  geom_boxplot() +
  coord_cartesian(ylim=c(0,250)) +
  xlab("Borough") +
  ylab("Price")
```



```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=bedrooms, y=price, group=bedrooms), data=airbnb) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0,1500), xlim = c(0,10)) +
  scale_x_continuous(breaks=seq(0,10,1)) +
  xlab("Number of Bedrooms") +
  ylab("Price")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=bedrooms, y=price), data=airbnb) +
  scale_y_log10() +
  geom_line(stat="summary", fun.y=median) +
  scale_x_continuous(breaks=seq(0,16,1)) +
  xlab("Number of Bedrooms") +
  ylab("Median Price")
```

The plots above support the claim that being located in a desirable borough 
(Manhattan or Brooklyn) and having a large number of bedrooms are indicative 
of an expensive listing.

With regard to the number of bedrooms, statistical calculations tell a similar 
story:

```{r echo=FALSE, message=FALSE, warning=FALSE}
cor.test(airbnb$bedrooms, airbnb$price)
```

The small p-value lends credibility to our argument that the number of bedrooms 
is a statistically significant predictor of the price of a listing. However, the 
R-squared value of 0.08 obtained suggests that the number of bedrooms alone is 
insufficient to form a predictive model.

From here, we can proceed with investigating the correlation between price and 
several other factors. Below is a plot of the availability of a listing (within 
the 365 days ending on 02 October, 2017) and its price.

We might reasonably hypothesize that listings that are used by owners to exploit
a short-term vacancy of their property could be listed at a different price than 
listings of property whose purpose is to generate income through Airbnb 
throughout the year.

In the following analysis, we only take into account listings that are active 
within the past year (available for booking in at least 1 day out of 365).

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=availability_365, y=price), data=subset(airbnb, 
                                                     availability_365 > 0)) +
  geom_jitter(alpha=1/4) +
  ylim(0, quantile(airbnb$price, 0.99)) +
  xlab('Number of Days Available') +
  ylab('Price')
```

There seems to be no discernible relationship between a listing's availability 
and price. It appears that listings which are exclusively used year-round for 
Airbnb rental purposes are no more or less expensive than listings that are put 
up simply to fill a temporary vacancy.

Another factor that might affect the price of listings is its review score. 
These are the scores assigned by visitors of any particular listing. A listing 
with high review scores will attract more prospective visitors and tenants, and 
as such it will be interesting to examine if hosts with highly-reviewed listings 
set a higher price to exploit the popularity and attractiveness of their homes.

We will only be examining homes with more than 10 reviews.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=review_score, y=price), 
       data=subset(airbnb, number_of_reviews > 10)) +
  geom_jitter(alpha=1/4) +
  ylim(0, quantile(airbnb$price, 0.99)) +
  xlab('Review Score') +
  ylab('Price')
```

We can see from the plot that there seems to be some evidence to support the 
claim that higher review scores relate to higher prices. A causal relationship 
cannot, however, be definitively proven.

Naturally, the number of guests included in a listing must have an impact on 
the price of a listing. The greater the number of guests a host is expected to 
cater, the higher the cost should be to the visitors. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=guests_included, y=price, group=guests_included), data=airbnb) + 
  geom_boxplot() + 
  ylim(0, quantile(airbnb$price, 0.99)) + 
  xlim(0, 12.5) +
  xlab('Guests Included') +
  ylab('Price')
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
cor.test(airbnb$price, airbnb$guests_included)
```

The boxplot above confirms our assertion that a greater number of guests 
included in any particular listing translates to a higher price.

Airbnb states that 70% of all visits end up with a review. This means that we 
can at least expect the trend for the number of reviews per month to be closely 
associated with the trend for visits per month. With this knowledge, we can 
examine if cheaper listings do indeed attract more bookings.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=reviews_per_month, y=price), data=subset(airbnb)) +
  geom_jitter(alpha=1/4) +
  ylim(c(0,quantile(airbnb$price, 0.99))) +
  xlab('Reviews per Month') +
  ylab('Price')
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
cor.test(airbnb$price, airbnb$reviews_per_month)
```

Although the scatterplot above suggests a trend, the calculations show a very 
weak correlation between price and the number of reviews per month a listing 
receives. This finding, surprisingly, means that listings with lower prices 
might not necessarily attract more bookings. This does not necessarily suggest 
that Airbnb users are price inelastic when it comes to paying for lodging while 
travelling. It is entirely plausible that the budget of Airbnb users is highly 
variable, and that the service caters to more than just travellers looking for 
low-cost alternatives to hotels.

Another relevant relationship that we can investigate is the one between the 
number of reviews per month and the review scores of listings. Essentially, we 
are asking the quesiton: Are more highly-rated listings attracting more 
visitors?

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=reviews_per_month, y=review_score), data=airbnb) +
  geom_jitter(alpha=1/4) +
  xlab('Reviews per Month') +
  ylab('Review Score')
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
cor.test(airbnb$reviews_per_month, airbnb$review_score)
```

The answer, contrary to expectations, is no - more highly reviewed listings does 
not necessarily attract more bookings. Again, although the plot initially 
suggests a positive relationship between the two features, calculations shows 
that the correlation is extremely weak. Surprisingly, the data shows us that 
listings with high or perfect review scores are not necessarily more frequently 
reviewed (and thus not more frequently visited) than listings with lower review 
scores. This begs the question: If price and reviews are not the main 
determinants of the number of bookings a listing receives, what is?

Moving forward, we explore the relationship between price and cancellation 
policy. A listing with a flexible cancellation policy might be expected to come 
with a price premium. A suitable analogy are prices for hotel rooms, where lower 
rates are often offered for bookings that are non-refundable.

```{r echo=FALSE, message=FALSE, warning=FALSE}
airbnb.price_by_cancel <- airbnb %>%
  filter(cancellation_policy %in% c("flexible", "moderate", "strict")) %>%
  group_by(cancellation_policy) %>%
  summarise(mean_price = mean(price),
           median_price = median(price),
           n = n())
head(airbnb.price_by_cancel)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=cancellation_policy, y=price), 
       data=subset(airbnb, cancellation_policy %in% 
                     c("flexible", "moderate", "strict"))) +
  geom_boxplot() +
  coord_cartesian(ylim=c(0,250)) +
  xlab('Cancellation Policy') + 
  ylab('Price')
```


The plot above shows that stricter cancellation policies does not necessarily 
translate to lower prices. In fact, the data suggests that the opposite might 
even be true. The median and mean prices for listings with a strict cancellation 
policy are considerably higher than those with a flexible or moderate 
cancellation policy.

From the scatterplot matrix illustrated earlier, there are several other 
interesting relationships to explore between a listing's features. One 
relationship that stands out is the one between a listing's availability 
and its cancellation policy.

```{r echo=FALSE, message=FALSE, warning=FALSE}

airbnb.availability_by_cancel <- airbnb %>%
  filter(availability_365 > 0 & cancellation_policy %in% 
           c("flexible", "moderate", "strict")) %>%
  group_by(cancellation_policy) %>%
  summarise(mean_availability = mean(availability_365),
           median_availability = median(availability_365),
           n = n())
head(airbnb.availability_by_cancel)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=cancellation_policy, y=availability_365), 
       data=subset(airbnb, availability_365 > 0 & cancellation_policy %in% 
                     c("flexible", "moderate", "strict"))) +
  geom_boxplot() +
  xlab('Cancellation Policy') +
  ylab('Number of Days Available')
```

We observe that listings with higher availability tend to have a stricter 
cancellation policy. The reasons behind this are not immediately obvious. One 
could reasonably suggest that for homes used exclusively to generate income for 
Airbnb, the lack of a strict cancellation policy could lead to volatile 
fluctuations in the hosts' expected income. As these hosts place greater 
reliance on the income from their Airbnb homes, they thus demand greater 
commitment on the visitors' part in exchange for the flexible schedule they 
offer. Note that we only considered listings that were available for at least 
one day within the past year.

Below, we proceed to examine the variations in availability for listings in the 
different boroughs.

```{r echo=FALSE, message=FALSE, warning=FALSE}

airbnb.availability_by_borough <- airbnb %>%
  filter(availability_365 > 0) %>%
  group_by(borough) %>%
  summarise(mean_availability = mean(availability_365),
           median_availability = median(availability_365),
           n = n())
head(airbnb.availability_by_borough)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=borough, y=availability_365), 
       data=subset(airbnb, availability_365 > 0)) +
  geom_boxplot() +
  ylab('Number of Days Available') +
  xlab('Borough')
```

We see that listings in Staten Island and the Bronx tend to be more widely 
available in terms of schedule, while listings in Brooklyn and Manhattan tend 
to be less frequently available for booking. On average, listings in Brooklyn 
and Manhattan were only available for 191 and 185 days respectively over the 
past year. This compares unfavorably to Staten Island, where the average is 250 
days.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

As expected, the location and size of each listing provides a good indication of 
an Airbnb's relative price. Listings in Manhattan and Brooklyn are indeed more 
expensive. We also see that listings with a greater number of bedrooms tend to 
be priced higher.

Moreover, we also found that a listing's review score, number of guests included 
and cancellation policy are also indicative of price in a specific direction. 
Higher review scores and a greater amount of guests included both relate to 
higher prices. Expensive listings also tend to have a stricter cancellation 
policy.


### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

There seems to be a relationship between a listing's availability and its 
cancellation policy. Listings that are available for more days in a year tend to
have stricter cancellation policies. This disparity is even greater if we take 
into account listings that have not been active (available for at least a day) 
in the 365 days ending on 02 October, 2017.

Another interesting observation is that a listing's popularity appears to be 
independent of its price. Cheaper listings does not equate to greater demand for 
those listings.

Also, great reviews do not seem to guarantee frequent visitors, as our analysis 
found that listings with high or perfect review scores do not attract 
significantly more bookings than listings with lower scores.

### What was the strongest relationship you found?

The strongest relationship found was the positive relationship between price and
the number of bedrooms.

# Multivariate Plots Section

In our following analysis concerning the number of bedrooms, we only focus on 
listings with 5 or less bedrooms, as this consists of the majority and prevents 
bigger and more expensive listings to skew our results in any way.

```{r echo=FALSE, Multivariate_Plots}
ggplot(aes(x=bedrooms, y=price), 
       data = subset(airbnb, room_type='Entire home/apt')) +
  geom_line(stat='summary', fun.y=median, 
            aes(color = borough)) +
  labs(color="Neighbourhood") +
  xlab('Number of Bedrooms') +
  ylab('Median Price') +
  xlim(c(0,5))
```

As extra bedrooms are only relevant for a listing that provides an entire home 
or apartment, we did not include listings offering only a shared or private room 
in this analysis.

The above plot lends support to our conclusion that listings in Manhattan and 
Brooklyn are generally more expensive than the listings in the remaining 
boroughs. For Manhattan, this is especially true for homes with a higher number 
of bedrooms.

The plot also illustrates that the price premium attached to listings in 
Manhattan are, unsurprisingly, the most significant. The gap between the price 
of Manhattan listings and those of other boroughs is evident, and this gap 
increases disproportionately as the number of bedrooms in the listing increases.

On the other hand, the premium on listings in Brooklyn is much less significant, 
as the median price of listings across all number of bedrooms is not far from 
that of listings in Queens.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=bedrooms, y=price), data = subset(airbnb, cancellation_policy %in% 
                                                 c('flexible', 'moderate', 
                                                      'strict'))) +
  geom_line(stat='summary', fun.y=median, aes(color = cancellation_policy)) +
  labs(color="Cancellation Policy") +
  xlab('Number of Bedrooms') +
  ylab('Median Price') +
  xlim(c(0,5))

ggplot(aes(x=bedrooms, y=price, group=bedrooms), data = subset(airbnb, cancellation_policy %in% 
                                                 c('flexible', 'moderate', 
                                                      'strict'))) +
  geom_boxplot() +
  xlab('Number of Bedrooms') +
  ylab('Median Price') +
  facet_wrap(~cancellation_policy) +
  scale_x_continuous(limits=c(-0.5,5.5), breaks=seq(0,5,1)) +
  ylim(c(0, quantile(airbnb$price, 0.95)))
```

The line chart above illustrates our discovery that listings with a stricter 
cancellation policy tend to be more expensive, regardless of the size of the 
listing.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=review_score, y=price), 
       data=subset(airbnb, number_of_reviews > 10 & 
                     price < quantile(airbnb$price, 0.99))) +
  geom_jitter(alpha=1/4) +
  facet_wrap(~borough, scales='free') +
  xlab('Review Score') + 
  ylab('Price')
```

From the plots above, we see that in all boroughs, most of the listings have 
review scores of between 80-100. Keep in mind that the plots above are generated 
for listings with more than 10 reviews and whose price is under the 99th 
percentile of listings in each borough.

This observation seems consistent over all boroughs, regardless of the price 
differences of listings between boroughs. The higher prices of listings in 
Manhattan and Brooklyn does not seem to impact this range, as expensive listings 
do not have signficantly better or worse scores than other listings. 

Earlier in the report, we posited that the impact of a listing's number of 
bedrooms differ greatly for listings that offer entire homes and listings that 
only offer private or shared rooms. 

Here, we examine the relationship between price and bedrooms given the room type 
provided by a particular listing. We will take into account only listings with 6 
or less bedrooms, as the listings offering private rooms have a maximum size of 
only 6 bedrooms. Furthermore, we will not consider shared rooms in our analysis 
as the sample size is relatively insignificant.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=bedrooms, y=price), data=subset(airbnb, bedrooms <= 6 & 
                                               room_type != 'Shared room')) +
  geom_line(aes(color=room_type), stat='summary', fun.y='median') +
  labs(color="Room Type") +
  xlab('Number of Bedrooms') +
  ylab('Price')
```

The relationship between the median price of listings and the number of bedrooms 
is completely unsimilar for entire homes and private rooms. There is an obvious 
upward trend for listings that offer the entire home, while the number of 
bedrooms does not seem to have any impact when the listing is only offering a 
single room for use.

Earlier, we plotted the number of bedrooms in a listing against its price and 
found that there seemed to be a rather significant positive correlation between 
the two features. However, when we consider the fact that analysis done on the 
number of bedrooms thus far includes listings that offer private rooms and 
shared rooms, we see now that our methodology is somewhat flawed. 

As we have demonstrated above, for private rooms and shared rooms, the number of 
bedrooms in the listing should be significantly less important than if the 
listing is for the entire home. Regardless of the listing's size, any 
prospective visitor will only be paying for a single room. As such, we would 
expect to arrive at more convincing figures if we exclude listings for private 
and shared rooms.

```{r echo=FALSE, message=FALSE, warning=FALSE}
m1 <- lm(formula = price ~ bedrooms, data = subset(airbnb, room_type == 
                                                     'Entire home/apt'))
mtable(m1)
```

The resultant numbers confirms our suspicion, and we do indeed obtain stronger 
evidence for the positive relationship between price and the number of bedrooms 
when we only take into account listings that offer an entire home or apartment.

Using what we have learnt so far, we attempt to build a mathematical model that 
might help in predicting the price of an Airbnb listing when given relevant 
features.

```{r echo=FALSE, message=FALSE, warning=FALSE}
m1 <- lm(formula = price ~ bedrooms, data=airbnb)
m2 <- update(m1, ~ . + review_score)
m3 <- update(m2, ~ . + guests_included)
m4 <- update(m3, ~ . + room_type)
m5 <- update(m4, ~ . + borough)
mtable(m1, m2, m3, m4, m5)
```

We were able to build a decent but insufficient model in predicting the price of 
Airbnb listings. Ultimately, I would assert that any quantitative attempt to 
form an accurate predictive model of Airbnb listing prices will be somewhat 
flawed as the most important indicator of price is the aesthetic of the home 
listed. Our willingness to pay is still largely influenced by our impression of 
the photographs of the listings. Listings with high quality photographs of a 
beautiful home will undoubtedly command a higher price. The quality of the 
photographs, however, is not a quantifiable feature that we can account for in 
our calculations. Furthermore, taste is a highly subjective matter.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

The significance of location in determining a listing's price is enhanced when 
we examined it across listings with different numbers of bedrooms. We notice 
that the price premium for listings in Manhattan increases with the number of 
bedrooms. This is unsurprising as homes with numerous bedrooms are especially 
difficult to find in Manhattan. We also found that the premium for listings in 
Brooklyn is not as significant as we might have previously thought.

Also, when considering only listings that offer an entire home or apartment, the 
correlation between price and the number of bedrooms grows signficantly 
stronger. This is because we have shown that there is little or no correlation 
between price and number of bedrooms for listings offering only private or 
shared rooms. This makes absolute sense as it hardly matters how large a home is 
if the visitor is only paying for and utilizing a single room.

### Were there any interesting or surprising interactions between features?

The observation that listings with a strict cancellation policy tend to be 
slightly more expensive than listings with a flexible or moderate cancellation 
policy is rather surprising, and goes against what we know about traditional 
lodging options such as hotels. In hotels, discounted rates are usually 
accompanied by strict cancellation policies where the amount paid is 
non-refundable. When it comes to Airbnb listings, however, we observe a 
contradictory trend but without a satisfactory explanation.

However, this apparent premium that we observe is negligible, and attempting to 
add a listing's cancellation policy into our model does not improve it much, or 
at all.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

We did attempt to create a model of Airbnb listing prices. However, our model 
does not seem to be sufficient or satisfactory in predicting the price of Airbnb 
listings given the relevant features. An R-squared value of 0.207 suggests that 
the features we have included are insufficient in providing a reliable model of 
Airbnb listing prices. An important limitation of our model is that listings 
that have not been rated by visitors were not taken into consideration, as 
review scores are an important feature of our model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(aes(x=price), data=airbnb) +
  geom_histogram(binwidth=0.05, fill=I('#FFA500')) +
  scale_x_log10() +
  xlab("Price") +
  ylab("Number of Listings") +
  ggtitle('Log_10 Airbnb Listing Prices')
```

### Description One

The overall log price distrbution of Airbnb listings in New York City seems to 
be fairly normal, although prices vary across many different factors, notably 
location, size and room type.

### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(aes(x=borough, y=availability_365), 
       data=subset(airbnb, availability_365 > 0)) +
  geom_boxplot() +
  xlab('Borough') +
  ylab('Number of Days Available') +
  ggtitle('Number of Days Available by Borough')
```

### Description Two

Brooklyn and especially Manhattan has the highest volume of Airbnb listings, and 
listings in these two boroughs are generally more expensive and in high demand. 
However, the plot above demonstrates that hosts of the listings in Brooklyn and 
Manhattan tend to make their homes available on Airbnb for considerably less 
days in a year compared to hosts in the Bronx, Queens and Staten Island.

### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(aes(x=bedrooms, y=price), data = subset(airbnb, cancellation_policy %in% 
                                                 c('flexible', 'moderate', 
                                                   'strict'))) +
  geom_line(stat='summary', fun.y=median, aes(color = cancellation_policy)) +
  ggtitle('Price by Number of Bedrooms and Cancellation Policy') +
  labs(color="Cancellation Policy") +
  xlab('Number of Bedrooms') +
  ylab('Median Price') +
  xlim(c(0,5))
```

### Description Three

The above is a plot of the relationship between a listing's price and its size, 
grouped by 3 types of cancellation policies: flexible, moderate or strict. It 
tells us that for the same number of bed rooms, cheaper listings tend to have a 
more flexible cancellation policy. Although the price premium attached to strict 
cancellation policies is not significant, it was a rather surprising discovery.

------

# Reflection

The Airbnb dataset contains information on more than 44,000 Airbnb listings in 
New York City which are spread across 96 variables. The very first challenge was 
to identify and isolate the variables that were relevant with what I was trying 
to achieve. The need to be concise was constantly confronted by the fear of 
missing out on something important. Working with such a large data set also 
necessitated constant revision of previous findings. Throughout the project, I 
constantly made discoveries that challenged initial observations and 
assumptions. Nevertheless, this led to more insightful evaluation of the 
relationships between certain features.

Among the features that we have studied, the following have rather significant 
correlations with price: the number of bedrooms, the borough in which the 
listing is located in, room type, guests included and the listing's review 
score. These features are able to provide a basic, albeit insufficient, 
indication of the price of a particular Airbnb listing. Some surprising 
relationships were discovered. For instance, a listing's review scores and 
availability had less of an impact on price than we expected. We also discovered 
that cheaper listings did not seem to attract a significantly greater amount of 
bookings. Lastly, we found that listings with strict cancellation policies tend 
to be more expensive than listings with flexible ones.

The original data set is massive, and we have only taken into account 11 the 
features we deemed interesting. However, there are many more features available 
for even more in-depth analysis of the data. For instance, listings have 
specific review scores for cleanliness, communication, location, value, etc. An 
interesting question to ask here is which of these rating criteria do users 
place the most emphasis on. Also, with data on how long the hosts have been 
active on Airbnb as well as the number of listings each host has, we can see if 
there are any trends that pertain more to 'experienced' hosts rather than 
'inexperienced' ones. 
